Bullet Data Preparation for LIME Diagnostics Paper

### 2020-09-28 code/02-data-preparation.Rmd
moving code to LIME-research-journals

@@

## Compare Dimensions to rtrees

The dimensions of the raw data:

```{r}
bullet_train_raw %>% dim()
```

The dimension of the data once the observations with tank rash (`flag != FALSE`) are removed:

```{r}
bullet_train_raw %>% filter(flag == FALSE) %>% dim()
```

The number of observations in the random forest `rtrees`:

```{r}
length(bulletxtrctr::rtrees$predicted)
```

The number of rows in the training data with the flagged observations excluded does not match the number of observations used to train `rtrees`. In the Hare, Hofmann, and Carriquiry (2017), they describe removing four land impressions that were flagged for quality assessment:

- Barrel 6 Bullet 2-1
- Barrel 9 Bullet 2-4
- Unknown Bullet B-2
- Unknown Bullet Q-4

It is unclear which study these lands correspond to. The code below removes all comparisons that include these barrel, bullet, and land combinations:

```{r}
bullet_train_filtered <- 
  bullet_train_raw %>%
  mutate(bbl1 = str_remove(land_id1, pattern = "Hamby252-|Hamby173-"),
         bbl2 = str_remove(land_id2, pattern = "Hamby252-|Hamby173-")) %>%
  filter(!(bbl1 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4") | 
           bbl2 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4"))) %>%
  select(-bbl1, -bbl2) 
```

The dimensions now agree with the number of observations in `rtrees`:

```{r}
dim(bullet_train_filtered)
length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_filtered)[1]
```

The code below removes observations based on Eric Hare's code (available [here](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R)), which he thinks may be the code used to train `rtrees`:

```{r}
bullet_train_eric_filter <- 
  bullet_train_raw %>%
  separate(col = "land_id1", into = c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(col = "land_id2", into = c("study2", "barrel2", "bullet2", "land2")) %>%
  filter(study1 != "Hamby173" | bullet1 != "BE") %>%
  filter(study2 != "Hamby173" | bullet2 != "BE")

# Eric's code for reference:
# filter(study.y != "Hamby44" | barrel.y != "E") %>%
# filter(study.x != "Hamby44" | barrel.x != "E") %>%
```

Here, the dimensions to not agree with the number of observations in `rtrees`:

```{r}
dim(bullet_train_eric_filter)
length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_eric_filter)[1]
```

Extracting the predictions on the training data from `rtrees` and using `rtrees` to make predictions on the filtered data:

```{r}
rtrees_pred = predict(bulletxtrctr::rtrees, type = "prob")[,2]

new_pred = predict(bulletxtrctr::rtrees, 
                   bullet_train_filtered %>% 
                     select(rownames(bulletxtrctr::rtrees$importance)), type = "prob")[,2]

new_pred_full = predict(bulletxtrctr::rtrees, 
                        bullet_train_raw %>%
                          select(rownames(bulletxtrctr::rtrees$importance)), type = "prob")[,2]
```

Plot comparing the two predictions (we discovered that the `predict` function from random forest returns different results when applied to only the model (out-of-bag predictions) and the training data (regular predictions), so this is not a good way to test for agreement between training data and models):

```{r}
data.frame(rtrees_pred = sort(rtrees_pred), new_pred = sort(new_pred)) %>%
  ggplot(aes(rtrees_pred, new_pred)) + 
  geom_point()
```

Plot showing distributions of predictions from `rtrees` and new predictions made on full training data (same comment here as above about not being able to use this to determine if we have the correct training data for the model):

```{r}
data.frame(data = c(rep("rtrees", length(rtrees_pred)), 
                    rep("new predictions on full data", length(new_pred_full))), pred = c(rtrees_pred, new_pred_full)) %>%
  ggplot(aes(x = pred, fill = data)) + 
  geom_density(alpha = 0.75)
```

Comparison to the data Eric Hare said was used to fit `rtrees` which has half the number of observations as the number of rtrees predictions:

```{r}
eric <- read.csv("https://raw.githubusercontent.com/erichare/imaging-paper/master/data/data-25-25/bullet-stats.csv")
dim(eric)
```

>>>>

Load the newest training data provided by Heike (on 9/24/2020 from CSAFE data base):

```{r}
ccfs_withlands_raw = read.csv("../data/raw/CCFs_withlands.csv")
```

@@
